---
title: "Model predictions for logistic regressions"
author: "oriana"
date: "February 17, 2015"
output:
  html_document:
    highlight: haddock
    theme: cerulean
---

This is document will demonstrate how to calculate and plot the predictions of a mixed-effects logistic regression. This based largely on Ben Bolker's Owls data example [[pdf]](http://glmm.wdfiles.com/local--files/examples/Owls.pdf), & help from [Morgan Sonderegger](people.linguistics.mcgill.ca/~morgan/). This document also greatly benefitted from the R Markdown Reference Guide [[pdf]](http://rmarkdown.rstudio.com/RMarkdownReferenceGuide.pdf).

First, we have to load the following libraries:
```{r message=FALSE,error=FALSE,warning=FALSE}
library(languageR)
library(lme4)
library(lmerTest)
library(optimx)
library(arm)
library(ggplot2)
```

* `languageR` contains the freely available data set we'll be using,
* `lme4` is what we need to fit the model,
* `ggplot2` for making nice plots. (Useful ggplot2 reference: [Cookbook for R](http://www.cookbook-r.com/Graphs/))

### Fitting a model

Let's fit a model with the following information:

* `CaseMarking` as the dependent variable: `1` corresponds to ergative marking, `0` to other (this is modified from the original data set) 
* Random effects:
	+ `Text`
	+ `Speaker`
* Fixed Effects:
	+ `WordOrder`
	+ `AgeGroup`
	+ `AnimacyOfSubject`

```{r echo=FALSE,message=FALSE}
warlpiri$CaseMarking <- factor(warlpiri$CaseMarking,levels=c('other','ergative'))
````

```{r cache=TRUE}
warlpiri.lmer = glmer(CaseMarking ~ WordOrder + AgeGroup +
                        AnimacyOfSubject + (1|Text) + (1|Speaker),
                      control=glmerControl(optimizer="optimx",optCtrl=list(method="nlminb")),
                      family = "binomial", data = warlpiri)

summary(warlpiri.lmer)


```


### Getting the predictions

Set up the frame that gives you all possible combinations of values. 

```{r}
pframe0 <- with(warlpiri, expand.grid(WordOrder=levels(WordOrder),AgeGroup=levels(AgeGroup),AnimacyOfSubject=levels(AnimacyOfSubject)))

pframe0
```

Now `pframe0` contains all possible values for a given observation. 

*(How do you do this if there's an interaction in your model?)*

And from that, you can get a matrix with all the different possible values that a given observation might have. 
```{r}
mm <- model.matrix(~WordOrder+AgeGroup+AnimacyOfSubject,data=pframe0)

pframe1 <- data.frame(pframe0,eta=mm%*%fixef(warlpiri.lmer))
pframe1 <- with(pframe1,data.frame(pframe1,CaseMarking=invlogit(eta)))

pframe1$pse <- diag(mm %*% tcrossprod(vcov(warlpiri.lmer), mm))
```
Which makes these the bounds of the 95% confidence interval for the average word/speaker prediction, in log-odds. 
```{r}
pframe1$hi <- with(pframe1, eta + 1.96*pse)
pframe1$low <- with(pframe1, eta - 1.96*pse)
```

Translate these numbers to probablity using the `invlogit()` function. 

```{r}
pframe1$hi.prob <- invlogit(pframe1$hi)
pframe1$low.prob <- invlogit(pframe1$low)
```

Now you can make a nice plot of your predicted probabilities with 95% confidence intervals!

```{r}
ggplot(aes(x=AgeGroup, ymin=low.prob, ymax=hi.prob,y=CaseMarking), data=pframe1) + geom_errorbar(aes(color=AnimacyOfSubject),width=0.2) + geom_point(aes(color=AnimacyOfSubject)) + coord_cartesian(ylim=c(0, 1)) + facet_wrap(~AnimacyOfSubject+WordOrder) + ggtitle('Predicted probability ')
```

```{r}
ggplot(aes(x=AnimacyOfSubject, ymin=low.prob, ymax=hi.prob,y=CaseMarking), data=pframe1) + geom_errorbar(aes(width=0.2,color=AgeGroup))+ coord_cartesian(ylim=c(0, 1)) + ggtitle('Predicted probability ') + facet_grid(~WordOrder)
```




